Generating Images From Captions\\ With Attention

### 2015-10-30 iclr-paper-new.tex
ugly version of the paper

In this paper, we illustrate how sequential deep learning techniques can be used to build a conditional probabilistic model over natural image space effectively. By extending the Deep Recurrent Attention Writer (DRAW) \citep{gregor_draw}, our model iteratively draws patches on a canvas, while attending to the relevant words in the description. Overall, the main contributions of this work are the following: we introduce a conditional alignDRAW model, a generative model of images from captions using a soft attention mechanism. The images generated by our alignDRAW model are refined in a post-processing step by a deterministic Laplacian pyramid adversarial network \citep{denton_lapgan}. We then illustrate how our method, learnt on Microsoft COCO \citep{mscoco}, generalizes to captions describing novel scenarios that are not seen in the dataset.

@@

@@

\newpage
\section{Experiments}
\subsection{MNIST With Captions}
As a first experiment, we trained our proposed model on the MNIST dataset with artificial captions. Either one or two digits from the MNIST training dataset were placed on a $60 \times 60$ blank image. One digit was placed in one of the four (top-left, top-right, bottom-left or bottom-right) corners of the image. Two digits were either placed horizontally or vertically in nonoverlapping fashion. 
While many generative models were trained on the binarized version of the MNIST dataset, we trained directly on pixel intensities with a binary cross-entropy cost function.

@@

\subsubsection{Analysis of Generated Images}
The main goal of this work is to learn a model that can understand the semantic meaning expressed in the textual descriptions of images, such as the properties of objects, the relationships between them, etc. and then use that knowledge to generate relevant images. To examine the understanding of our model, we wrote a set of captions inspired by the COCO dataset and changed some words in the captions to see whether the model made the relevant changes in the generated samples.

@@

\begin{figure}[!h]
\captionsetup[subfigure]{labelformat=empty}
\vspace{-0.3cm}
\begin{center}
\subfloat[A \underline{yellow} school bus parked in a parking lot.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-yellow-school-bus-parked-in-a-parking-lot-sharp.png}}\quad
%
\subfloat[A \underline{red} school bus parked in a parking lot.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-red-school-bus-parked-in-a-parking-lot-sharp.png}}\quad
%
\subfloat[A \underline{green} school bus parked in a parking lot.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-green-school-bus-parked-in-a-parking-lot-sharp.png}}\quad
%
\subfloat[A \underline{blue} school bus parked in a parking lot.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-blue-school-bus-parked-in-a-parking-lot-sharp.png}}\quad
%
\end{center}
\caption{Examples of changing the color while keeping the caption fixed.}
\label{fig:genimages1}
\vspace{-0.3cm}
\end{figure}

\begin{figure}[!t]
\captionsetup[subfigure]{labelformat=empty}
%\vspace{-0.5cm}
\begin{center}
\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-very-large-commercial-plane-flying-in-blue-skies-closest.png}}\quad
%
\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-very-large-commercial-plane-flying-in-rainy-skies-closest.png}}\quad
%
\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-herd-of-elephants-walking-across-a-dry-grass-field-closest.png}}\quad
%
\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-herd-of-elephants-walking-across-a-green-grass-field-closest.png}}\\
\vspace{-0.45cm}
%
\subfloat[A very large commercial plane flying in \underline{blue} skies.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-very-large-commercial-plane-flying-in-blue-skies-sharp.png}}\quad
%
\subfloat[A very large commercial plane flying in \underline{rainy} skies.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-very-large-commercial-plane-flying-in-rainy-skies-sharp.png}}\quad
%
\subfloat[A herd of elephants walking across a \underline{dry} grass field.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-herd-of-elephants-walking-across-a-dry-grass-field-sharp.png}}\quad
%
\subfloat[A herd of elephants walking across a \underline{green} grass field.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-herd-of-elephants-walking-across-a-green-grass-field-sharp.png}}\quad
%
\end{center}
\caption{\textbf{Bottom}: Examples of changing the background while keeping the caption fixed. \textbf{Top}: The respective nearest training images based on pixel-wise L2 distance. The nearest images from the training set also indicate that the model was not simply copying the patterns it observed during the learning phase.}
\label{fig:genimages2}
\vspace{-0.3cm}
\end{figure}

@@

\begin{figure}[!h]
\captionsetup[subfigure]{labelformat=empty}
\vspace{-0.3cm}
\begin{center}
\subfloat[\underline{The decadent chocolate} \underline{desert} is on the table.
]{\includegraphics[width=0.23\textwidth]{figures/new/the-decadent-chocolate-dessert-is-on-the-table-sharp.png}}\quad
%
\subfloat[\underline{A bowl of bananas} is on the table.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-bowl-of-bananas-is-on-the-table-sharp.png}}\quad
%
\subfloat[A vintage photo of a \underline{cat}.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-vintage-photo-of-a-cat-sharp.png}}\quad
%
\subfloat[A vintage photo of a \underline{dog}.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-vintage-photo-of-a-dog-sharp.png}}\quad
%
\end{center}
\caption{Examples of changing the object while keeping the caption fixed.}
\label{fig:genimages3}
\vspace{-0.3cm}
\end{figure}

@@

\begin{figure}[!h]
\captionsetup[subfigure]{labelformat=empty}
\vspace{-0.3cm}
\begin{center}
\subfloat[A stop sign is flying in blue skies.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-stop-sign-is-flying-in-blue-skies-sharp.png}}\quad
%
\subfloat[A herd of elephants flying in the blue skies.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-herd-of-elephants-flying-in-the-blue-skies-sharp.png}}\quad
%
\subfloat[A toilet seat sits open in the grass field.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-toilet-seat-sits-open-in-the-grass-field-sharp.png}}\quad
%
\subfloat[A person skiing on sand clad vast desert.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-person-skiing-on-sand-clad-vast-desert-sharp.png}}\quad
%
\end{center}
\caption{Examples of novel scene compositions.}
\label{fig:genimages4}
\vspace{-0.3cm}
\end{figure}

@@

\begin{figure}[!h]
\captionsetup[subfigure]{labelformat=empty}
\vspace{-0.3cm}
\begin{center}
\subfloat[\hlcthree{A} rider \hlcone{on} a blue \hlcone{motorcycle} in the \underline{\hlctwo{desert}}.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-rider-on-a-blue-motorcycle-in-the-desert-sharp.png}}\quad
%
\subfloat[\hlcthree{A} rider \hlcone{on} a blue \hlcone{motorcycle} in the \underline{\hlctwo{forest}}.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-rider-on-a-blue-motorcycle-in-the-forest-sharp.png}}\quad
%
\subfloat[\hlctwo{A} \hlcone{surfer}, a woman, and a child walk on the \underline{\hlctwo{beach}}.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-surfer-,-a-woman-,-and-a-child-walk-on-the-beach-sharp.png}}\quad
%
\subfloat[\hlcthree{A} \hlcone{surfer}, a woman, and a child walk on the \underline{\hlczero{sun}}.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-surfer-,-a-woman-,-and-a-child-walk-on-the-sun-sharp.png}}\quad
%
\end{center}
\caption{Examples of most attended words while changing the background in the caption.}
\label{fig:attentimages1}
\vspace{-0.3cm}
\end{figure}

@@

\begin{figure}[!h]
\captionsetup[subfigure]{labelformat=empty}
\vspace{-0.3cm}
\begin{center}
\subfloat[alignDRAW]{\includegraphics[width=0.23\textwidth]{figures/new/a-group-of-people-walk-on-a-beach-with-surf-boards-sharp-small.png}}\quad
%
\subfloat[LAPGAN]{\includegraphics[width=0.23\textwidth]{figures/a-group-of-people-walk-on-a-beach-with-surf-boards-lapgan-small.png}}\quad
%
\subfloat[Conv-Deconv VAE]{\includegraphics[width=0.23\textwidth]{figures/a-group-of-people-walk-on-a-beach-with-surf-boards-convdeconvvae-small.png}}\quad
%
\subfloat[Fully-Conn VAE]{\includegraphics[width=0.23\textwidth]{figures/a-group-of-people-walk-on-a-beach-with-surf-boards-fcvae-small.png}}\quad
%
\end{center}
\caption{Four different models displaying results from sampling caption \textit{A group of people walk on a beach with surf boards.}}
\label{fig:diffmodels}
\vspace{-0.3cm}
\end{figure}

@@

\newpage
\section{Acknowledgements}
We would like to thank developers of Theano \citep{theano} for a powerful software and authors of \citep{denton_lapgan} for open sourcing their code. We would also like to thank Ryan Kiros and Nitish Srivastava for helpful discussions. We will be releasing our code with pretrained models as soon as we clean it.

>>>>

In this paper, we illustrate how sequential deep learning techniques can be used to build a conditional probabilistic model over natural image space effectively. By extending the Deep Recurrent Attention Writer (DRAW) \citep{gregor_draw}, our model iteratively draws patches on a canvas, while attending to the relevant words in the description. Overall, the main contributions of this work are the following: we introduce a conditional alignDRAW model, a generative model of images from captions using a soft attention mechanism. The images generated by our alignDRAW model are refined in a post-processing step by a deterministic Laplacian pyramid adversarial network \citep{denton_lapgan}. We then illustrate how our method, learnt on Microsoft COCO \citep{mscoco}, generalizes to captions describing novel scenarios that are not seen in the dataset (Figure~\ref{fig:genimages4}).

@@

\begin{figure}[!t]
\captionsetup[subfigure]{labelformat=empty}
\vspace{-0.3cm}
\begin{center}
\subfloat[A stop sign is flying in blue skies.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-stop-sign-is-flying-in-blue-skies-sharp.png}}\quad
%
\subfloat[A herd of elephants flying in the blue skies.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-herd-of-elephants-flying-in-the-blue-skies-sharp.png}}\quad
%
\subfloat[A toilet seat sits open in the grass field.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-toilet-seat-sits-open-in-the-grass-field-sharp.png}}\quad
%
\subfloat[A person skiing on sand clad vast desert.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-person-skiing-on-sand-clad-vast-desert-sharp.png}}\quad
%
\end{center}
\caption{Examples of novel scene compositions.}
\label{fig:genimages4}
\vspace{-0.3cm}
\end{figure}

@@

%\newpage
\section{Experiments}
\subsection{MNIST With Captions}
As a first experiment, we trained our proposed model on the MNIST dataset with artificial captions. Either one or two digits from the MNIST training dataset were placed on a $60 \times 60$ blank image. One digit was placed in one of the four (top-left, top-right, bottom-left or bottom-right) corners of the image. Two digits were either placed horizontally or vertically in nonoverlapping fashion. 
While many generative models were trained on the binarized version of the MNIST dataset, we trained directly on pixel intensities with a binary cross-entropy cost function.

@@

\subsubsection{Analysis of Generated Images}

The main goal of this work is to learn a model that can understand the semantic meaning expressed in the textual descriptions of images, such as the properties of objects, the relationships between them, etc. and then use that knowledge to generate relevant images. To examine the understanding of our model, we wrote a set of captions inspired by the COCO dataset and changed some words in the captions to see whether the model made the relevant changes in the generated samples.

@@

\begin{figure}[!t]
\captionsetup[subfigure]{labelformat=empty}
\vspace{-0.1cm}
\begin{center}
\subfloat[A \underline{yellow} school bus parked in a parking lot.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-yellow-school-bus-parked-in-a-parking-lot-sharp.png}}\quad
%
\subfloat[A \underline{red} school bus parked in a parking lot.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-red-school-bus-parked-in-a-parking-lot-sharp.png}}\quad
%
\subfloat[A \underline{green} school bus parked in a parking lot.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-green-school-bus-parked-in-a-parking-lot-sharp.png}}\quad
%
\subfloat[A \underline{blue} school bus parked in a parking lot.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-blue-school-bus-parked-in-a-parking-lot-sharp.png}}\quad
%
\end{center}
\caption{Examples of changing the color while keeping the caption fixed.}
\label{fig:genimages1}
\vspace{-0.1cm}
\end{figure}

\begin{figure}[!t]
\captionsetup[subfigure]{labelformat=empty}
\vspace{-0.1cm}
\begin{center}
\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-very-large-commercial-plane-flying-in-blue-skies-closest.png}}\quad
%
\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-very-large-commercial-plane-flying-in-rainy-skies-closest.png}}\quad
%
\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-herd-of-elephants-walking-across-a-dry-grass-field-closest.png}}\quad
%
\subfloat[]{\includegraphics[width=0.23\textwidth]{figures/new/a-herd-of-elephants-walking-across-a-green-grass-field-closest.png}}\\
\vspace{-0.45cm}
%
\subfloat[A very large commercial plane flying in \underline{blue} skies.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-very-large-commercial-plane-flying-in-blue-skies-sharp.png}}\quad
%
\subfloat[A very large commercial plane flying in \underline{rainy} skies.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-very-large-commercial-plane-flying-in-rainy-skies-sharp.png}}\quad
%
\subfloat[A herd of elephants walking across a \underline{dry} grass field.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-herd-of-elephants-walking-across-a-dry-grass-field-sharp.png}}\quad
%
\subfloat[A herd of elephants walking across a \underline{green} grass field.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-herd-of-elephants-walking-across-a-green-grass-field-sharp.png}}\quad
%
\end{center}
\caption{\textbf{Bottom}: Examples of changing the background while keeping the caption fixed. \textbf{Top}: The respective nearest training images based on pixel-wise L2 distance. The nearest images from the training set also indicate that the model was not simply copying the patterns it observed during the learning phase.}
\label{fig:genimages2}
\vspace{-0.1cm}
\end{figure}

\begin{figure}[!t]
\captionsetup[subfigure]{labelformat=empty}
\vspace{-0.1cm}
\begin{center}
\subfloat[\underline{The decadent chocolate} \underline{desert} is on the table.
]{\includegraphics[width=0.23\textwidth]{figures/new/the-decadent-chocolate-dessert-is-on-the-table-sharp.png}}\quad
%
\subfloat[\underline{A bowl of bananas} is on the table.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-bowl-of-bananas-is-on-the-table-sharp.png}}\quad
%
\subfloat[A vintage photo of a \underline{cat}.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-vintage-photo-of-a-cat-sharp.png}}\quad
%
\subfloat[A vintage photo of a \underline{dog}.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-vintage-photo-of-a-dog-sharp.png}}\quad
%
\end{center}
\caption{Examples of changing the object while keeping the caption fixed.}
\label{fig:genimages3}
\vspace{-0.1cm}
\end{figure}

\begin{figure}[!t]
\captionsetup[subfigure]{labelformat=empty}
\vspace{-0.1cm}
\begin{center}
\subfloat[\hlcthree{A} rider \hlcone{on} a blue \hlcone{motorcycle} in the \underline{\hlctwo{desert}}.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-rider-on-a-blue-motorcycle-in-the-desert-sharp.png}}\quad
%
\subfloat[\hlcthree{A} rider \hlcone{on} a blue \hlcone{motorcycle} in the \underline{\hlctwo{forest}}.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-rider-on-a-blue-motorcycle-in-the-forest-sharp.png}}\quad
%
\subfloat[\hlctwo{A} \hlcone{surfer}, a woman, and a child walk on the \underline{\hlctwo{beach}}.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-surfer-,-a-woman-,-and-a-child-walk-on-the-beach-sharp.png}}\quad
%
\subfloat[\hlcthree{A} \hlcone{surfer}, a woman, and a child walk on the \underline{\hlczero{sun}}.
]{\includegraphics[width=0.23\textwidth]{figures/new/a-surfer-,-a-woman-,-and-a-child-walk-on-the-sun-sharp.png}}\quad
%
\end{center}
\caption{Examples of most attended words while changing the background in the caption.}
\label{fig:attentimages1}
\vspace{-0.1cm}
\end{figure}

\begin{figure}[!h]
\captionsetup[subfigure]{labelformat=empty}
\vspace{-0.1cm}
\begin{center}
\subfloat[alignDRAW]{\includegraphics[width=0.23\textwidth]{figures/new/a-group-of-people-walk-on-a-beach-with-surf-boards-sharp-small.png}}\quad
%
\subfloat[LAPGAN]{\includegraphics[width=0.23\textwidth]{figures/a-group-of-people-walk-on-a-beach-with-surf-boards-lapgan-small.png}}\quad
%
\subfloat[Conv-Deconv VAE]{\includegraphics[width=0.23\textwidth]{figures/a-group-of-people-walk-on-a-beach-with-surf-boards-convdeconvvae-small.png}}\quad
%
\subfloat[Fully-Conn VAE]{\includegraphics[width=0.23\textwidth]{figures/a-group-of-people-walk-on-a-beach-with-surf-boards-fcvae-small.png}}\quad
%
\end{center}
\caption{Four different models displaying results from sampling caption \textit{A group of people walk on a beach with surf boards.}}
\label{fig:diffmodels}
\vspace{-0.1cm}
\end{figure}

@@

@@

@@

@@

@@

%\newpage
\section{Acknowledgements}
We would like to thank developers of Theano \citep{theano} for a powerful software and authors of \citep{denton_lapgan} for open sourcing their code. We would also like to thank Ryan Kiros and Nitish Srivastava for helpful discussions. We will be releasing our code with pretrained models as soon as we clean it.
